Random Forest to find the top important features

### 2019-11-01 scripts/151_RF_find_top_features_other.Rmd
update

```{r}
# the species table from metaphlan2
# the species abundance here are actually in percent
s_tb <- read_tsv('../data/metaphlan2_result/merged_abundance_table_species.txt') %>% 
  rename_all(funs(str_replace(., '_profile$','')))

@@

### Use more abundant features(species abundance > 1%) + uncorrelated pheno var as predictors

@@

```{r}
# to use the top N percent features

# to test on different threshold of selecting most abundant species
sum_thre <- seq(0.5, 1, 0.05)
tree_num <- 1000

res <- sum_thre %>% 
  set_names(sum_thre) %>% 
  map(function(thre) {
    s_sum_top <- quantile(s_sum, thre)
    combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > s_sum_top])], pheno %>% 
                    select(Toxicity, ORR, Disease_Status_CAR))

    tibble(
      num_fea = length(s_sum[s_sum > s_sum_top]) + 2,
      OOB_err_rate = randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T , ntree = tree_num )[['err.rate']][tree_num, 1]

    ) 
  }) %>% 
  bind_rows(.id = 'thre')

res %>% 
  kable(caption = 'Top N fraction features and the OOB error rate') %>% 
  kable_styling(full_width = F, position = 'left')
```

### See the important features(species)

@@

```{r}
# repeat the RF M = 100 times
# pull top 10 features
N <- 10 
M <- 100 
pull_top_features <- function(){
  importance(randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T ) , type=1, scale = F) %>% 
    as.data.frame() %>% 
    rownames_to_column(var = 'species') %>% 
    arrange(desc(MeanDecreaseAccuracy)) %>% 
    top_n(N) %>% 
    pull(species)
}

@@

stable_top_fea %>% 
  kable(caption = 'Frequency of features shown as top 10 in 100 random forest runs') %>% 
  kable_styling(full_width = F, position = 'left')

@@

What's interesting is this species *Ruminococcus gnavus*. [This recent paper](https://www.pnas.org/content/116/26/12672) points out that *Ruminococcus gnavus* produces a polysaccharide molecule that induces the production of inflammatory cytokines.

And [this paper](https://www.ncbi.nlm.nih.gov/pubmed/29330424) mentions that "*Faecalibacterium prausnitzii* decrease and *Ruminococcus gnavus* increase were the essential characteristics in CHF patients' gut microbiota".

>>>>

```{r}
# the species table from metaphlan2
# the species abundance here are actually in percent
s_tb <- read_tsv('../data/metaphlan2_result/merged_abundance_table_phylum.txt') %>% 
  rename_all(funs(str_replace(., '_profile$','')))

@@

### Use more abundant features(phylum abundance > 1%) + uncorrelated pheno var as predictors

@@

### See the important features(phylum)

@@

```{r}
# repeat the RF M = 100 times
# pull top 10 features
N <- 3 
M <- 100 
pull_top_features <- function(){
  importance(randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T ) , type=1, scale = F) %>% 
    as.data.frame() %>% 
    rownames_to_column(var = 'species') %>% 
    arrange(desc(MeanDecreaseAccuracy)) %>% 
    top_n(N) %>% 
    pull(species)
}

@@

stable_top_fea %>% 
  kable(caption = 'Frequency of features shown as top 3 in 100 random forest runs') %>% 
  kable_styling(full_width = F, position = 'left')

@@

### 2019-11-01 scripts/16_RF_find_top_pathways.Rmd
update

```{r}
# the species table from metaphlan2
# the species abundance here are actually in percent
s_tb <- read_tsv('../data/metaphlan2_result/merged_abundance_table_species.txt') %>% 
  rename_all(funs(str_replace(., '_profile$','')))

s_tb_sum <- s_tb %>% 
  column_to_rownames('ID') %>% 
  as.matrix() %>% 
  apply(., 2, sum)

@@

# to put the feature table in the sample as rowname format to do the ML 
s_tb_rf <- s_tb %>% 
  column_to_rownames('ID') %>% 
  as.matrix() %>% 
  t

@@

impor_feature <- importance(tox_RF, type=1) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = 'species') %>% 
  arrange(MeanDecreaseAccuracy)

@@

>>>>

```{r}
# the species table from metaphlan2
# the species abundance here are actually in percent
s_tb <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
  rename(pathway = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  # remove the unmapped and the unintegrated record
  filter(!pathway %in% c('UNMAPPED','UNINTEGRATED')) %>% 
  mutate(pathway = str_extract(pathway, '^.+:')) %>% 
  mutate(pathway = str_replace_all(pathway, ':','')) %>% 
  mutate(pathway = str_replace_all(pathway, '-','_'))%>% 
  mutate(pathway = str_replace_all(pathway, '\\+','_'))

@@

# to put the feature table in the sample as rowname format to do the ML 
s_tb_rf <- s_tb %>% 
  column_to_rownames('pathway') %>% 
  as.matrix() %>% 
  t

@@

@@

### Use top N fraction features + uncorrelated pheno var as predictors

### 2019-11-01 scripts/17_RF_CR_find_top_features.Rmd
update

---
title: "Random Forest to find the top important features"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(randomForest)
library(ggpubr)
library(tidyverse)
library(kableExtra)
library(arrangements)
library(caret)
```

@@

There are much for Y patients than N. If all assigning to Y it will probably like this.

### Use all features + all pheno var as predictors

```{r} 
# bind the pheno with the transposed cts matrix
combined <- cbind(s_tb_rf, pheno %>% 
                    select(-Sampleid))

@@

combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > 1])], pheno %>% 
                    select(-Sampleid))

@@

No improvements at all. I think because the relatively small sample size and uneven distribution of the groups.

@@

res <- sum_thre %>% 
  set_names(sum_thre) %>% 
  map(function(thre) {
    s_sum_top <- quantile(s_sum, thre)
    combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > s_sum_top])], pheno %>% 
                    select(-Sampleid))

@@

    ) 
  }) %>% 
  bind_rows(.id = 'thre')
```

Interesting, using no feature info actually has smaller OOB error rate.

@@

### See the important features(species)

@@

```{r}
# repeat the RF M = 100 times
# pull top 10 features
N <- 10
M <- 100 
pull_top_features <- function(){
  importance(randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T ) , type=1, scale = F) %>% 
    as.data.frame() %>% 
    rownames_to_column(var = 'species') %>% 
    arrange(desc(MeanDecreaseAccuracy)) %>% 
    top_n(N) %>% 
    pull(species)
}

@@

What's interesting is this species *Ruminococcus gnavus*. [This recent paper](https://www.pnas.org/content/116/26/12672) points out that *Ruminococcus gnavus* produces a polysaccharide molecule that induces the production of inflammatory cytokines, which could be hypothesized to cause toxicity too.

>>>>

---
title: "Random Forest to find the top important features"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(randomForest)
library(ggpubr)
library(tidyverse)
library(kableExtra)
library(arrangements)
```

@@

### Use all features + uncorrelated pheno var as predictors

uncorrelated pheno var: Disease_Status_CAR for CR

```{r} 
# bind the pheno with the transposed cts matrix
combined <- cbind(s_tb_rf, pheno %>% 
                    select(Disease_Status_CAR, CR))

@@

combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > 1])], pheno %>% 
                    select(Disease_Status_CAR, CR))

@@

### Use top N fraction features + uncorrelated pheno var as predictors

@@

res <- sum_thre %>% 
  set_names(sum_thre) %>% 
  map(function(thre) {
    s_sum_top <- quantile(s_sum, thre)
    combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > s_sum_top])], pheno %>% 
                    select(Disease_Status_CAR, CR))

@@

    ) 
  }) %>% 
  bind_rows(.id = 'thre')

res %>% 
  kable(caption = 'Top N fraction features and the OOB error rate') %>% 
  kable_styling(full_width = F, position = 'left')
```

@@

### See the important features (species)

@@

```{r}
# repeat the RF M = 100 times
# pull top 10 features
N <- 10
M <- 100 
pull_top_features <- function(){
  importance(randomForest(formula = CR ~ ., data = combined_fil,  importance = T ) , type=1, scale = F) %>% 
    as.data.frame() %>% 
    rownames_to_column(var = 'species') %>% 
    arrange(desc(MeanDecreaseAccuracy)) %>% 
    top_n(N) %>% 
    pull(species)
}

@@
